---
title: "Final Version of MyProject_PSY329"
author: "Mursal Jahed"
date: "2024-10-11"
output: "github_document"
---
Load packages
```{r}
library(haven)
library(dplyr)
library(ggplot2)
library(tidyr) 
library(psych)
library(car)
library(lmtest)
library(performance)
library(sjPlot)
library(bruceR)
```

Load dataset
```{r}
load("/Users/mursal_j/Downloads/ICPSR_36561/DS0001/36561-0001-Data.rda")
```

Select variables
```{r}
list_projectdata <- da36561.0001 %>% 
  select(PARCIVT1, PARCIV_1T1, PARCIV_4T1, PARCIV_2T1, PARCIV_3T1, COMMMEANT1, YIIACTS_3T1, YIIACTS_5T1, YIIACTS_11T1, YIIACTS_13T1, YIIACTS_15T1, PEERINVT1, PEERCIV_1T1, PEERCIV_2T1, PEERCIV_3T1, GENDERT1, GRADET1, CRITNEWST1, CRITNEWS_1T1, CRITNEWS_2T1, AGE_YEART1)

#rename variables for easier interpretation & conciseness
list_projectdata <- list_projectdata %>%
  rename(
    Teen_Civic_Engagement = COMMMEANT1,
    Parents = PARCIVT1,
    Peers = PEERINVT1,
    Critical_Consumption = CRITNEWST1
  )

colnames(list_projectdata)

```
Demographics
```{r}
mean_age <- mean(list_projectdata$AGE_YEART1, na.rm = TRUE)
print(mean_age)
gender_counts <- table(list_projectdata$GENDERT1)
print(gender_counts)
gender_percent <- prop.table(gender_counts) * 100
print(gender_percent)
```

Checking for Assumptions
```{r}
model <- lm(Teen_Civic_Engagement ~ Parents + Peers, data = list_projectdata)
check_model(model)
summary(model)
```
Analysis - How do parental vs peer civic involvement influence civic involvement in teenagers?
```{r}
plot_model(model,  type ="est",  show.values = TRUE, vline.color = "#1B191999", line.size = 1.5, dot.size = 2.5, colors = "blue")+ theme_classic() + ggtitle("Parental and Peer Influence on Teen Civic Engagement") +
  xlab("Predictors") +  # Label for x-axis
  ylab("Estimates")
tab_model(model)
```

Analysis - How does this relationship change if the teens' critical news consumption is taken into account?
Checking assumptions
```{r}
model <- lm(Teen_Civic_Engagement ~ Critical_Consumption + Parents + Peers, data = list_projectdata)
check_model(model)
summary(model)
```
Analysis
```{r}
plot_model(model, type = "est", show.values = TRUE, vline.color = "#1B191999", 
           line.size = 1.5, dot.size = 2.5, colors = "blue", 
           terms = c("Critical_Consumption", "Parents", "Peers")) + theme_classic() + ggtitle("Critial News Consumption Influences Teen Civic Involvement") +
  xlab("Predictors") +  # Label for x-axis
  ylab("Estimates")
tab_model(model)
```

Testing reliability
```{r}
#Reliability of Teen_Civic_Engagement/Teen community service
Alpha(list_projectdata, vars=c("YIIACTS_3T1", "YIIACTS_5T1", "YIIACTS_11T1", "YIIACTS_13T1", "YIIACTS_15T1"))
#Reliability of Peers/Peer civic involvement
Alpha(list_projectdata, vars=c("PEERCIV_1T1", "PEERCIV_2T1", "PEERCIV_3T1"))
#Reliability of Parents/Parental civic involvement
Alpha(list_projectdata, vars=c("PARCIV_1T1", "PARCIV_4T1", "PARCIV_2T1", "PARCIV_3T1"))
```
Reliability of Critical_Consumption (Spearman's Rho)
```{r}
# need to create a data set and only select those two
spearman_for_Critical_Consumption <- da36561.0001 %>% 
  select(CRITNEWS_1T1, CRITNEWS_2T1)
Corr(spearman_for_Critical_Consumption, method = "spearman")
```


Factor Analysis
```{r}
#FA of COMMMEANT1/Teen community service
EFA(list_projectdata, vars=c("YIIACTS_3T1", "YIIACTS_5T1", "YIIACTS_11T1", "YIIACTS_13T1", "YIIACTS_15T1"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```

```{r}
#FA of PEERCIVACT1/Peer civic involvement
EFA(list_projectdata, vars=c("PEERCIV_1T1", "PEERCIV_2T1", "PEERCIV_3T1"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))

```

```{r}
#FA of PARCIVT1/Parental civic involvement
EFA(list_projectdata, vars=c("PARCIV_1T1", "PARCIV_2T1", "PARCIV_3T1", "PARCIV_4T1"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))
```

